---
title: "Assignment 06: Airbnb Market Analysis with dplyr"
author: "Jack Hibbard"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(pillar.sigfig = 6)
```

## Read the data

```{r read-data}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/midwest_airbnb_extended.db")
listings     = DBI::dbReadTable(con, "listings") |> tibble::as_tibble()
availability = DBI::dbReadTable(con, "availability_monthly") |> tibble::as_tibble()
reviews      = DBI::dbReadTable(con, "reviews_monthly") |> tibble::as_tibble()
DBI::dbDisconnect(con)

dplyr::glimpse(listings)
```

## Task 1

Who owns the market: listings per `host_id` across all three cities, joined back to the listings.

```{r task1}
host_counts = listings |>
  dplyr::count(host_id, name = "host_listings")

listings_hosts = listings |>
  dplyr::left_join(host_counts, by = "host_id") |>
  dplyr::mutate(big_host = host_listings >= 10)

# A1: hosts with 10 or more listings
a1 = host_counts |>
  dplyr::filter(host_listings >= 10) |>
  dplyr::summarize(hosts_10_plus = dplyr::n())
a1

# A2 (and B2): percent of each city's listings held by those hosts
task1_share = listings_hosts |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    listings = dplyr::n(),
    big_host_listings = sum(big_host),
    pct_big_host = 100 * mean(big_host)
  )
task1_share
```

## Task 2

Value for a family of four: well-reviewed entire homes that sleep at least 4.

```{r task2}
family_value = listings |>
  dplyr::filter(
    room_type == "Entire home/apt",
    accommodates >= 4,
    number_of_reviews >= 10,
    review_scores_rating >= 4.8
  ) |>
  dplyr::mutate(price_per_person = price / accommodates) |>
  dplyr::group_by(city, neighbourhood) |>
  dplyr::summarize(
    n_listings = dplyr::n(),
    median_ppp = median(price_per_person),
    .groups = "drop"
  ) |>
  dplyr::filter(n_listings >= 20)

# A3: Columbus neighbourhood with the lowest median price per person
family_value |>
  dplyr::filter(city == "Columbus") |>
  dplyr::arrange(median_ppp)
```

## Task 3

What earns: medians by city and room type in one `summarize()` call with `dplyr::across()`.

```{r task3}
earnings = listings |>
  dplyr::group_by(city, room_type) |>
  dplyr::summarize(
    dplyr::across(
      c(price, reviews_per_month, estimated_revenue_l365d),
      ~ median(.x, na.rm = TRUE),
      .names = "median_{.col}"
    ),
    n_listings = dplyr::n(),
    .groups = "drop"
  )
earnings

# A4: median estimated revenue of Twin Cities entire homes
earnings |>
  dplyr::filter(city == "Twin Cities", room_type == "Entire home/apt") |>
  dplyr::select(city, room_type, median_estimated_revenue_l365d)
```

## Task 4

Concentration: share of each city's estimated revenue earned by its top 10 percent of listings.

```{r task4}
task4_share = listings |>
  dplyr::group_by(city) |>
  dplyr::arrange(dplyr::desc(estimated_revenue_l365d), .by_group = TRUE) |>
  dplyr::mutate(
    rank = dplyr::row_number(),
    top10 = rank <= ceiling(0.10 * dplyr::n())
  ) |>
  dplyr::summarize(
    listings = dplyr::n(),
    top10_listings = sum(top10),
    total_revenue = sum(estimated_revenue_l365d, na.rm = TRUE),
    top10_revenue = sum(estimated_revenue_l365d[top10], na.rm = TRUE),
    pct_revenue_top10 = 100 * top10_revenue / total_revenue
  )

# A5: Columbus row
task4_share
```

## Task 5

The year ahead: October 2026 availability for entire homes.

```{r task5}
october = listings |>
  dplyr::filter(room_type == "Entire home/apt") |>
  dplyr::select(id, city) |>
  dplyr::inner_join(
    availability |> dplyr::filter(month == "2026-10"),
    by = "id"
  ) |>
  dplyr::mutate(mostly_taken = nights_available < 0.5 * nights_in_month)

# A6 and A7: city with the lowest share of mostly-taken entire homes
october |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    entire_homes = dplyr::n(),
    share_mostly_taken = mean(mostly_taken),
    n_mostly_taken = sum(mostly_taken)
  ) |>
  dplyr::mutate(pct_mostly_taken = 100 * share_mostly_taken) |>
  dplyr::arrange(share_mostly_taken)
```

## Task 6

Demand by season: reviews per city and month, July 2025 through June 2026.

```{r task6}
monthly_demand = reviews |>
  dplyr::inner_join(listings |> dplyr::select(id, city), by = "id") |>
  dplyr::filter(month >= "2025-07", month <= "2026-06") |>
  dplyr::group_by(city, month) |>
  dplyr::summarize(total_reviews = sum(reviews), .groups = "drop")

# A8 and A9: Twin Cities busiest month and busiest-to-quietest ratio
twin = monthly_demand |>
  dplyr::filter(city == "Twin Cities") |>
  dplyr::arrange(dplyr::desc(total_reviews))
twin

twin |>
  dplyr::summarize(
    busiest_month = month[which.max(total_reviews)],
    busiest = max(total_reviews),
    quietest_month = month[which.min(total_reviews)],
    quietest = min(total_reviews),
    ratio = busiest / quietest
  )

# All three cities for comparison
monthly_demand |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    busiest_month = month[which.max(total_reviews)],
    quietest_month = month[which.min(total_reviews)],
    ratio = max(total_reviews) / min(total_reviews)
  )
```

## Task 7

The Superhost premium among entire homes with a known Superhost flag.

```{r task7}
superhost = listings |>
  dplyr::filter(
    room_type == "Entire home/apt",
    host_is_superhost %in% c("t", "f")
  ) |>
  dplyr::group_by(city, host_is_superhost) |>
  dplyr::summarize(
    n_listings = dplyr::n(),
    median_price = median(price, na.rm = TRUE),
    median_rating = median(review_scores_rating, na.rm = TRUE),
    .groups = "drop"
  )
superhost

# Premium = superhost median price minus other-host median price (dollars)
superhost |>
  dplyr::group_by(city) |>
  dplyr::summarize(
    superhost_median_price = median_price[host_is_superhost == "t"],
    other_median_price = median_price[host_is_superhost == "f"],
    premium = superhost_median_price - other_median_price,
    superhost_median_rating = median_rating[host_is_superhost == "t"],
    other_median_rating = median_rating[host_is_superhost == "f"]
  )
```

## Task 8

What the join lost: listings with no rows in `availability`.

```{r task8}
no_calendar = listings |>
  dplyr::anti_join(availability, by = "id")

# A10
no_calendar |>
  dplyr::summarize(listings_without_availability = dplyr::n())

# Which city loses the most
no_calendar |>
  dplyr::count(city, name = "lost_listings") |>
  dplyr::arrange(dplyr::desc(lost_listings))
```

## B2 support: both shares for all three cities

```{r b2}
task1_share |>
  dplyr::select(city, pct_big_host) |>
  dplyr::inner_join(task4_share |> dplyr::select(city, pct_revenue_top10), by = "city")
```
